---
title: "CEMA Technical Interview"
author: "Anita Makori"
date: "2023-03-15"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load packages
library(tidyverse)
library(janitor)
library(sf)
```


```{r, include=F, echo=FALSE, message=FALSE, warning=FALSE}
# 1. Data Wrangling Challenge

# read data & clean names
household_data <- read_csv("L4H_household_baseline_sample.csv", na="---")%>%
  clean_names()

individual_data <- read_csv("L4H_individual_baseline_sample.csv", na="---")%>%
  clean_names()

mother_data <- read_csv("L4H_mother_baseline_sample.csv", na="---")%>%
  clean_names()

# filter eligible hhs
household_data1 <- household_data%>%
  filter(hh_eligible%in%"1")

# merge datasets
mother_individual_merged <- individual_data%>%
  full_join(mother_data, by=c("number"="number_0"))

household_merged_data <- household_data%>%
  full_join(mother_individual_merged, by="household_id")

# recode variables
household_merged_data1 <- household_merged_data%>%
  mutate_at(vars(reason_for_ineligibility, rspntgndr, h_hfrml_eductn, rspndtmarital, rspndt_edctn, maincme), funs(factor))%>%
  mutate(reason_for_ineligibility=recode(reason_for_ineligibility, "1"="No adult occupier >16 years", "2"="Withdrawal", "3"="Other reason"))%>%
  mutate(rspntgndr=recode(rspntgndr, "1"="Male", "2"="Female"))%>%
  mutate(h_hfrml_eductn=recode(h_hfrml_eductn, "1"="Not completed Primary school", "2"="Primary school", "3"="Secondary school", "4"="College-graduate", "5"="Madrassa", "6"="Other"))%>%
  mutate(rspndtmarital=recode(rspndtmarital, "1"="Single", "2"="Married monogamous", "3"="Married polygamous", "4"="Divorced/ separated", "5"="Widow(er)"))%>%
  mutate(rspndt_edctn=recode(rspndt_edctn, "1"="No formal education", "2"="Primary School", "3"="Secondary school", "4"="College-graduate", "5"="Madrassa", "6"="Other"))%>%
  mutate(maincme=recode(maincme, "1"="Sale of livestock & livestock products", "2"="Sale of crops", "3"="Trading/business", "4"="Employment (salaried income)", "5"="Sale of personal assets", "6"="Remittance", "7"="Other"))

# separate
household_merged_data2 <- household_merged_data1%>%
  separate(lvstckown, c("lvstckown1", "lvstckown2", "lvstckown3", "lvstckown4", "lvstckown5", "lvstckown6", "lvstckown7", "lvstckown8", "lvstckown9", "lvstckown10", "lvstckown11", "lvstckown12", "lvstckown13", "lvstckown14", "lvstckown15"), " ")%>%
  separate(herdynamics, c("herdynamics1", "herdynamics2", "herdynamics3", "herdynamics4", "herdynamics5", "herdynamics6", "herdynamics7"), " ")

# new column
household_merged_data3 <- household_merged_data2%>%
  mutate(study_arm = ifelse(village.x%in%c("Lependera", "Saale-Sambakah", "Namarei", "Manyatta Lengima", "Lokoshula", "TubchaDakhane", "Rengumo-Gargule"), "Study arm 1", ifelse(  village.x%in%c("Uyam  village", "Rongumo_kurkum", "Manyatta K.A.G", "Ltepes Ooodo", "Lbaarok1"), "Study arm 2", "Study arm 3")))

# new object
herd_dynamics <- household_merged_data3%>%
  select(interview_date.x, household_id, study_arm, cwsbrth, shpbrth, goatsbrth, cmlsbrth, calves_death, bulls_death, cows_death, sheep_death, msheep_death, fsheep_death, goats_death, mgoats_death, fgoats_death, camels_death, mcamels_death, fcamels_death, cowsgft, sheepgfts, goatsgft, cmlsgft, cowsgvnout, sheepgvnout, goatsgvnout, cmlsgvnout)

# new column
herd_dynamics1 <- herd_dynamics%>%
  separate(interview_date.x, c("year", "month", "day"), "-")%>%
  mutate(monthyear=paste0(year, "-", month))

#
herd_dynamics2 <- herd_dynamics1%>%
  group_by(study_arm, monthyear)%>%
  mutate(cow_births=sum(cwsbrth, na.rm = T))%>%
  mutate(shp_births=sum(shpbrth, na.rm = T))%>%
  mutate(gt_births=sum(goatsbrth, na.rm = T))%>%
  mutate(cml_births=sum(cmlsbrth, na.rm = T))%>%
  mutate(cow_deaths=sum(c(calves_death, bulls_death, cows_death), na.rm = T))%>%
  mutate(shp_deaths=sum(c(sheep_death, msheep_death, fsheep_death), na.rm = T))%>%
  mutate(gt_deaths=sum(c(goats_death, mgoats_death, fgoats_death), na.rm = T))%>%
  mutate(cml_deaths=sum(c(camels_death, mcamels_death, fcamels_death), na.rm = T))%>%
  mutate(cow_gifts=sum(cowsgft, na.rm = T))%>%
  mutate(shp_gifts=sum(sheepgfts, na.rm = T))%>%
  mutate(gt_gifts=sum(goatsgft, na.rm = T))%>%
  mutate(cml_gifts=sum(cmlsgft, na.rm = T))%>%
  mutate(cow_givenout=sum(cowsgvnout, na.rm = T))%>%
  mutate(shp_givenout=sum(sheepgvnout, na.rm = T))%>%
  mutate(gt_givenout=sum(goatsgvnout, na.rm = T))%>%
  mutate(cml_givenout=sum(cmlsgvnout, na.rm = T))%>%
  ungroup()

#subset
herd_dynamics3 <- herd_dynamics2%>%
  select(study_arm, monthyear, cow_births, shp_births, gt_births, cml_births, cow_deaths, shp_deaths, gt_deaths, cml_deaths, cow_gifts, shp_gifts, gt_gifts, cml_gifts, cow_givenout, shp_givenout, gt_givenout, cml_givenout)%>%
  distinct(study_arm, monthyear, .keep_all=T)
```

# 2. Data Visualization Challenge
```{r, echo=FALSE, fig.width=14, fig.height=8, message=FALSE, warning=FALSE}
#prep data
herd_dynamics4 <- herd_dynamics3%>%
  pivot_longer(cols = c(cow_births:cml_givenout), names_to = "category", values_to = "number")%>%
  mutate(dynamics=ifelse(grepl("births", category), "Births", ifelse(grepl("deaths", category), "Deaths", ifelse(grepl("gifts", category), "Gifts", "Given Out"))))%>%
  mutate(species=ifelse(grepl("cow", category), "Cows", ifelse(grepl("shp", category), "Sheep", ifelse(grepl( "gt", category), "Goats", "Camels"))))

# visuals 1
herd_dynamics_plot <- ggplot(herd_dynamics4,aes(x=monthyear, y=number, fill=species))+
  geom_col(position="dodge", stat="identity")+
  theme_bw()+
  facet_grid(rows=vars(dynamics))+
  labs( y="Number of animals", title="Herd Dynamics", x="Time period (year-month)", fill="Species")+
  scale_fill_manual(values = c("#d7191c","#fdae61","#ffffbf","#2c7bb6"))+
  theme(text=element_text(size=12))
herd_dynamics_plot

# read & prep data for visuals 2
teen_preg_data <- read_csv("table6_teenpregnancybycounty.csv")%>%
  mutate(County=recode(County,"Tharaka-Nithi"="Tharaka Nithi", "Elgeyo/Marakwet"="Elgeyo Marakwet","Murang’a"="Muranga", "Nairobi City"="Nairobi", "Taita/Taveta"="Taita Taveta"))

county <- st_read("County.shp")%>%
  rename(County=Name)

teen_preg_data1 <- full_join(county, teen_preg_data, by="County")

# visuals 2
teen_preg_map <- ggplot(teen_preg_data1)+
  geom_sf(data=teen_preg_data1, aes(geometry=geometry), fill=NA)+
  geom_sf(aes(fill=Ever_pregnant),color="grey80", size=0.0)+
  theme_void()+
  scale_fill_gradient(low = "white", high = "#8c2d04")+
  labs(x="", y="",fill="Proportion (%)", title = "Percentage of teenagers\n who have ever\n been pregnant by county")+
  theme(text=element_text(size=14))
teen_preg_map
```

# 3. Data Analytics Challenge
```{r, include=F, echo=FALSE, message=FALSE, warning=FALSE}
# read data
ideal_data <- read_csv("ideal3a.csv")%>%
  clean_names()%>%
  mutate(calf_sex=recode(calf_sex, "1"="Male", "2"="Female"))

#prep data for model
ideal_data$reasons_loss1 <- fct_relevel(ideal_data$reasons_loss1, "survived")
ideal_data$calf_sex <- fct_relevel(ideal_data$calf_sex, "Female")
ideal_data$distance_water <- fct_relevel(ideal_data$distance_water, "1-5 km")

# logistic regression model
ideal_glm_model <- glm(reasons_loss1~distance_water, data=ideal_data, family="binomial")
exp(cbind(OR=coef(ideal_glm_model),confint(ideal_glm_model)))

#calf_sex Male 8.38095111 3.25781115 28.53486053
#distance_water <1 km        0.2462571 0.1233357 0.4816015
#distance_water At household 0.4076923 0.1100863 1.2121723

# linear regression model
ideal_lm_model<-lm(adwg~manual_pcv, data = ideal_data)
exp(cbind(cc= coef(ideal_lm_model), confint(ideal_lm_model)))

#recruit_weight 1.004626 1.002877 1.006378
#manual_pcv  1.001705 1.000831 1.002580
```

### Factors associated with calves' likelihood of death

- Calf sex had a significant association with the likelihood of death with male calves having 8 times higher odds of death than female calves (OR=8.38, 95% CI=3.26,28.53)
- Proximity to the water point also had a significant association with calves' likelihood of death. The likelihood of death was reduced by 25% among calves from households within less than 1km from the water point in comparison with calves from household within 1-5km from the water point (OR=0.25, 95% CI=0.12, 0.48). 

### Factors associated with calves' growth rate

- Calves' weight during recruitment had a positive association with their growth rate. Higher weight at recruitment led to a higher growth rate (CC=1.005, 95% CI=1.003,1.006)

- The calves' health , as indicated by the Manual PCV, also had a positive association with their growth rate. A high manual PCV value led to a higher growth rate (CC=1.002, 95%CI=1.001,1.003)
